---
title: "STAT423-HW1"
author: "Nan Tang"
date: "1/17/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1-a

```{r message=FALSE, warning=FALSE}
library(alr4)
age_dt <- unique(wblake$Age)
leng_mean_dt <- numeric(length(age_dt))
for (i in age_dt) {
  leng_mean_dt[i] <- mean(wblake$Length[wblake$Age == i])
}
age_leng_mean_df <- data.frame(age_dt, leng_mean_dt)
plot(age_leng_mean_df, pch=16, ylab='Mean Length', xlab='Age', main='Age vs Mean of Length')
```

Base on the scatterplot, there is no strong evidence suggesting that linear relationshop exists between mean length for each age and age. 

## Question 1-b


```{r question-1-b}
leng_sd_dt <- numeric(length(age_dt))
for (i in age_dt) {
  leng_sd_dt[i] <- sd(wblake$Length[wblake$Age == i])
}
age_leng_sd_df <- data.frame(age_dt, leng_sd_dt)
plot(age_leng_sd_df, pch=16, ylab='SD of Length', xlab='Age', main='Age vs SD of Length')
```

The variance of standard deviation of length does not seem constant within ages. 

## Question 1-c

```{r question-1-c}
lm1 <- lm(wblake$Length~wblake$Scale)

plot(wblake$Scale, wblake$Length, pch=16, cex=0.5,
     xlab='Scale', ylab='Length' )
abline(lm1, col=2)

summary(lm1)

```

A simple linear regression line fits the data very well when scale is low, however, when scale is high, the residual increases. 

Base on the summary of linear regression, P-value of both intercept and coefficient of scale are much less than zero, indicating that both beta zero and beta one are unlikely to be zero. 


## Question 1-d

``` {r question-1-d}
resid1 <- summary(lm1)$residual
hist(resid1, xlab='Residuals', main='Histogram of Residuals', xlim=c(-100, 100))
qqnorm(resid1)
plot(lm(wblake$Length~wblake$Scale), which = 1)
```

The histogram shows the distribution of residuals is approximately normal. 
The linear pattern in qqplot also implies that the normality assumption of error is not violated. 
The pattern that shown in TA plot does not have a particular shape, indicating that normality assumption is not violate as well. 





